\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-21 doc/USER_MANUAL/04_creating_databases.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.6\textwidth]{figures/workflow.pdf}
\par\end{centering}

\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
specific model parameters.}

\label{fig:workflow.databases}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xgenerate_databases
\end{verbatim}
}
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

@@

Before running \texttt{xgenerate\_databases}, a number of parameters
need to be set in the main parameter \texttt{Par\_file} located in
the subdirectory \texttt{DATA}:
\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
 \texttt{PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
one slice of the whole mesh.
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
controls the length of the numerical simulation, i.e., twice the number
of time steps requires twice as much CPU time. This feature is not
used at the time of generating the distributed databases but is required
for the solver, i.e., you may change this parameter after running
\texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. This feature
is not used at the time of generating the distributed databases but
is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
for further details.
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use \texttt{Gmsh} \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.

\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.
To turn that option on, set parameter NUMBER\_OF\_SIMULTANEOUS\_RUNS to a value greater than 1.
To implement that, we create NUMBER\_OF\_SIMULTANEOUS\_RUNS MPI sub-communicators,
each of them being labeled "my\_local\_mpi\_comm\_world", and we use them
in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
we need to kill the entire run.
When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of NUMBER\_OF\_SIMULTANEOUS\_RUNS,
all the individual runs must use the same number of processor cores,
which as usual is NPROC in the Par\_file,
and thus the total number of processor cores to request from the batch system
should be NUMBER\_OF\_SIMULTANEOUS\_RUNS * NPROC.
All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
(with exactly four digits).

@@

The option NUMBER\_OF\_SIMULTANEOUS\_RUNS implements 3/.

\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}]:
if we perform simultaneous runs in parallel, if only the source and receivers vary between these runs
but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
(by a factor equal to NUMBER\_OF\_SIMULTANEOUS\_RUNS), and reducing I/Os is crucial in the case of huge runs.
Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
and the code will broadcast it to the others).

\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}]:
if one or a few of these simultaneous runs fail, kill all the runs or let the others finish using a fail-safe mechanism
(in most cases, should be set to true).

TODO / future work to do: currently the BROADCAST\_SAME\_MESH\_AND\_MODEL option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories, e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run. This would not be needed for forward simulations.

TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is. Thus these tools should work, but in their current version will need to be run for each simulation result independently.
More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters). This works because these routines consider multiple runs by design. You simply have to provide them the directories where the kernels are.

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim} 
where\\
rho\_s = solid density\\
rho\_f = fluid density\\
phi = porosity\\
tort = tortuosity\\
kxx = xx component of permeability tensor\\
kxy = xy,yx components of permeability tensor\\
kyy = yy component of permeability tensor\\
kxz = xz,zx components of permeability tensor\\
kzz = zz component of permeability tensor\\
kappa\_s = solid bulk modulus\\
kappa\_f= fluid bulk modulus\\
kappa\_fr= frame bulk modulus\\
eta\_f = fluid viscosity\\
mu\_fr = frame shear modulus.\\

>>>>

\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.6\textwidth]{figures/workflow.jpg}
\par
\end{centering}
\caption{Schematic workflow for a SPECFEM3D Cartesian simulation. The executable
\texttt{xgenerate\_databases} creates the GLL mesh points and assigns
specific model parameters.}
\label{fig:workflow.databases}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xgenerate_databases
\end{verbatim}
}
\noindent
Input for the program is provided through the main parameter file
\texttt{Par\_file}, which resides in the subdirectory \texttt{DATA}.
Please note that \texttt{xgenerate\_databases} must be called directly
from the main directory, as most of the binaries of the package.

@@

Before running \texttt{xgenerate\_databases}, a number of parameters
need to be set in the main parameter \texttt{Par\_file} located in
the subdirectory \texttt{DATA}:
\begin{description}
\item [{\texttt{SIMULATION\_TYPE}}] is set to 1 for forward simulations,
2 for adjoint simulations (see Section \ref{sec:Adjoint-simulation-finite})
and 3 for kernel simulations (see Section \ref{sec:Finite-Frequency-Kernels}).
\item [{\texttt{SAVE\_FORWARD}}] is only set to \texttt{.true.} for a forward
simulation with the last frame of the simulation saved, as part of
the finite-frequency kernel calculations (see Section \ref{sec:Finite-Frequency-Kernels}).
For a regular forward simulation, leave \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD} at their default values.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in the geographical coordinates,
and needs to be \texttt{.true.} when your model is specified in a
cartesian coordinates. \noun{UTM projection zone in which your simulation
region resides.}
\item [{\texttt{NPROC}}] The number of MPI processors, each one is assigned
one slice of the whole mesh.
\item [{\texttt{NSTEP}}] The number of time steps of the simulation. This
controls the length of the numerical simulation, i.e., twice the number
of time steps requires twice as much CPU time. This feature is not
used at the time of generating the distributed databases but is required
for the solver, i.e., you may change this parameter after running
\texttt{xgenerate\_databases}.
\item [{\texttt{DT}}] The length of each time step in seconds. This feature
is not used at the time of generating the distributed databases but
is required for the solver. Please see also Section~\ref{sec:Choosing-the-Time-Step}
for further details.
\item [{\texttt{NGNOD}}] The number of nodes for 2D and 3D shape functions
for hexahedra. We use either 8-node mesh elements (bricks) or 27-node
elements. If you use the internal mesher, the only option is 8-node
bricks (27-node elements are not supported). \texttt{CUBIT} does not
support HEX27 elements either (it can generate them, but they are
flat, i.e. identical to HEX8). To generate HEX27 elements with curvature
properly taken into account, you can use Gmsh \url{http://geuz.org/gmsh/}
\item [{\texttt{MODEL}}] Must be set to one of the following:

@@

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{src/shared/constants.h.in} and running the \texttt{configure} script again and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source. Originally,
if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
%
\item [{\texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}}] adds the ability to run several calculations (several earthquakes)
in an embarrassingly-parallel fashion from within the same run;
this can be useful when using a very large supercomputer to compute
many earthquakes in a catalog, in which case it can be better from
a batch job submission point of view to start fewer and much larger jobs,
each of them computing several earthquakes in parallel.

To turn that option on, set parameter \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} to a value greater than 1.
To implement that, we create \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} MPI sub-communicators,
each of them being labeled \texttt{my\_local\_mpi\_comm\_world}, and we use them
in all the routines in "src/shared/parallel.f90", except in MPI\_ABORT() because in that case
we need to kill the entire run.

When that option is on, of course the number of processor cores used to start
the code in the batch system must be a multiple of \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS},
all the individual runs must use the same number of processor cores,
which as usual is NPROC in the Par\_file,
and thus the total number of processor cores to request from the batch system
should be \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}~*~\texttt{NPROC}.
All the runs to perform must be placed in directories called run0001, run0002, run0003 and so on
(with exactly four digits).

@@

The option \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS} implements 3/.

\item [{\texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}}] if we perform simultaneous runs in parallel, 
if only the source and receivers vary between these runs
but not the mesh nor the model (velocity and density) then we can also read the mesh and model files
from a single run in the beginning and broadcast them to all the others; for a large number of simultaneous
runs for instance when solving inverse problems iteratively this can DRASTICALLY reduce I/Os to disk in the solver
(by a factor equal to \texttt{NUMBER\_OF\_SIMULTANEOUS\_RUNS}), and reducing I/Os is crucial in the case of huge runs.
Thus, always set this option to .true. if the mesh and the model are the same for all simultaneous runs.
In that case there is no need to duplicate the mesh and model file database (the content of the DATABASES\_MPI
directories) in each of the run0001, run0002,... directories, it is sufficient to have one in run0001
and the code will broadcast it to the others).

\item [{\texttt{USE\_FAILSAFE\_MECHANISM}}] if one or a few of these simultaneous runs fail, 
kill all the runs or let the others finish using a fail-safe mechanism
(in most cases, should be set to true).

TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL} 
option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. 
However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories,
e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run.
This would not be needed for forward simulations.

TODO / future work to do: the sensitivity kernel summing and smoothing tools in directory src/tomography are currently not ported to this new option to do many runs simultaneously, only the solver (src/specfem3d) is. Thus these tools should work, but in their current version will need to be run for each simulation result independently.

More precisely, the current kernel summing and smoothing routines work fine, with the exception that you need to move out the mesh files (and also the parameters). This works because these routines consider multiple runs by design. You simply have to provide them the directories where the kernels are.

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}
where\\
\texttt{rho\_s} = solid density,\\
\texttt{rho\_f} = fluid density,\\
\texttt{phi} = porosity,\\
\texttt{tort} = tortuosity,\\
\texttt{kxx} = xx component of permeability tensor,\\
\texttt{kxy} = xy,yx components of permeability tensor,\\
\texttt{kyy} = yy component of permeability tensor,\\
\texttt{kxz} = xz,zx components of permeability tensor,\\
\texttt{kzz} = zz component of permeability tensor,\\
\texttt{kappa\_s} = solid bulk modulus,\\
\texttt{kappa\_f} = fluid bulk modulus,\\
\texttt{kappa\_fr} = frame bulk modulus,\\
\texttt{eta\_f} = fluid viscosity,\\
\texttt{mu\_fr} = frame shear modulus.\\

### 2015-05-21 doc/USER_MANUAL/05_running_the_solver.tex
Buildbot ran the code beautifier.
updates manual for better wiki display

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
{\small
\begin{verbatim}
make xspecfem3D
\end{verbatim}
}
Please note that \texttt{xspecfem3D} must be called directly from
the main directory, as most of the binaries of the package.\\

\noindent
The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{}option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
 \texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\noindent
\begin{centering}
{\small \includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.pdf}
}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

@@

where $t_{\mathrm{pde}}$ is the time given in the first line of the
\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
is the time in the first column of the output seismogram.

\end{itemize}
\begin{figure}[htp]
\noindent \begin{centering}
\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod.pdf}
\par\end{centering}

\caption{Comparison of the shape of a triangle and the Gaussian function actually
used.}

\label{fig:gauss.vs.triangle}
\end{figure}

@@

\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics[width=5in]{figures/source_timing.pdf} }
\par\end{centering}

\caption{Example of timing for three sources. The center of the first source
triangle is defined to be time zero. Note that this is NOT in general
the hypocentral time, or the start time of the source (marked as tstart).
The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}

{\small \label{fig:source_timing} }
\end{figure}

The solver can calculate seismograms at any number of stations for
basically the same numerical cost, so the user is encouraged to include
as many stations as conceivably useful in the \texttt{STATIONS} file,
which looks like this:

\begin{figure}[H]
\noindent \begin{centering}
{\small \includegraphics{figures/STATIONS_basin_explained.pdf} }
\par\end{centering}

\label{fig:Sample-STATIONS-file.}
\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
be provided in geographical coordinates. The width of the station
label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
in the \texttt{constants.h} file), and the network label should be
no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
in the \texttt{constants.h} file).}
\end{figure}

>>>>

Now that you have successfully generated the databases, you are ready
to compile the solver. In the main directory, type
{\small
\begin{verbatim}
make xspecfem3D
\end{verbatim}
}
\noindent
Please note that \texttt{xspecfem3D} must be called directly from
the main directory, as most of the binaries of the package.\\

\noindent
The solver needs three input files in the \texttt{DATA} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed
in detail in the previous Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake
source parameter file or the force source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

@@

where $t_{\mathrm{pde}}$ is the time given in the first line of the
\texttt{CMTSOLUTION}, \texttt{time shift} is the corresponding value
from the original \texttt{CMTSOLUTION} file and $t_{\mathrm{synthetic}}$
is the time in the first column of the output seismogram.
\end{itemize}
%
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=3.5in]{figures/gauss_vs_triangle_mod.jpg}
\par
\end{centering}
\caption{Comparison of the shape of a triangle and the Gaussian function actually used.}
\label{fig:gauss.vs.triangle}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[width=4in]{figures/source_timing.jpg}
\par
\end{centering}
\caption{Example of timing for three sources. The center of the first source
triangle is defined to be time zero. Note that this is NOT in general
the hypocentral time, or the start time of the source (marked as tstart).
The parameter \texttt{time shift} in the \texttt{CMTSOLUTION} file
would be t1(=0), t2, t3 in this case, and the parameter \texttt{half
duration} would be hdur1, hdur2, hdur3 for the sources 1, 2, 3 respectively.}
{\small \label{fig:source_timing} }
\end{figure}

The solver can calculate seismograms at any number of stations for
basically the same numerical cost, so the user is encouraged to include
as many stations as conceivably useful in the \texttt{STATIONS} file,
which looks like this:
%
\begin{figure}[H]
\begin{centering}
\includegraphics[width=4in]{figures/STATIONS_basin_explained.jpg}
\par
\end{centering}
\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
be provided in geographical coordinates. The width of the station
label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
in the \texttt{constants.h} file), and the network label should be
no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
in the \texttt{constants.h} file).}
\label{fig:Sample-STATIONS-file.}
\end{figure}
